# Install R, RStudio, and R packages

R is the programming language and RStudio is the user interface to let us use R.

If you want to explore by yourself, here are the manuals to install [[**R**]{.underline}](https://cran.r-project.org/mirrors.html) and [[**RStudio**]{.underline}](https://posit.co/download/rstudio-desktop/).

In brief:

## Windows:

1.  Download the current R version for windows here: [[**R download link (R for windows)**]{.underline}](https://cran.rstudio.com/bin/windows/base/R-4.4.2-win.exe) and follow the instructions.

2.  Download RStudio here: [[**RStudio download link (windows)**]{.underline}](https://download1.rstudio.org/electron/windows/RStudio-2024.12.0-467.exe) and follow the instructions.

## macOS and Linux:

1.  Download the current R version for windows here: [**R for macOS (M1,2..Chip)**](https://cran.rstudio.com/bin/macosx/big-sur-arm64/base/R-4.4.2-arm64.pkg) \| [**R for macOS (Intel)**](https://cran.rstudio.com/bin/macosx/big-sur-x86_64/base/R-4.4.2-x86_64.pkg)[\| [**R for Ubuntu(Linux)**](https://cran.rstudio.com/bin/linux/ubuntu)]{.underline} and follow the instructions.
2.  Download RStudio here: [**RStudio for macOS**](https://download1.rstudio.org/electron/macos/RStudio-2024.12.0-467.dmg), [**,[RStudio for Ubuntu](https://download1.rstudio.org/electron/focal/amd64/rstudio-2024.04.1-748-amd64.deb)**]{.underline} and follow the instructions.

after installing R and RStudio, you can test it by opening the RStudio software and type a simple code in the console like:

Next step is to install all pre-requested packages for the course.

## Install pre-requested packages

Please type the below command in your console one by one:

```{r eval=FALSE, warning=FALSE}
# Step1. Install packages from CRAN
install.packages(c("dplyr", "Seurat", "devtools", "rmarkdown", "ggplot2", "ggpubr"))

# Step2. Install presto
devtools::install_github("immunogenomics/presto")

# Step3. Check if you have BiocManager
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

# Step4. Install hdf5r, DESeq2, edgeR
BiocManager::install(c("hdf5r", "DESeq2", "edgeR"))
```

## Check everything is settle

Check you have R and all packages are available for the **correct version**. Try execute the code below in your console in RStudio. You should get something like the content below without error. Try to install the same or later versions of the packages.

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(Seurat)
library(presto)
library(hdf5r)
library(DESeq2)
library(edgeR)
library(ggplot2)
library(ggpubr)

packages <- c("dplyr", "Seurat", "presto", "hdf5r", 
              "DESeq2", "edgeR", "ggplot2","ggpubr")
versions <- sapply(packages, function(pkg) paste0(
    pkg, " version: ", packageVersion(pkg)))
cat(versions, sep = "\n")

```

## 
